---
title: "time vs. fert"
author: "Emily Parsons"
date: "9/26/2022"
output: html_document
---

1. Age vs. Fertilization - all replicates no corrections

i. 2019 Night 1

```{r}
library(lattice)
rm(list=ls())
dat <- read.csv("C:/Users/parsonsee/Desktop/r_data/ofav1time.csv")
xyplot(fert~time,data=dat,type=c("p","r"), main = "OFAV 2019 Night 1: Time vs. Fert - SQ corrected avg (n=3)", xlab="Time Ova Added to Sperm", ylab="Fert Avg (n=3): Square Corrected")
m <- lm(fert_sq~time,data=dat)
anova(m)
```
```{r}
summary(m)
```




ii. 2019 Night 2

```{r}
library(lattice)
rm(list=ls())
dat <- read.csv("C:/Users/parsonsee/Desktop/ofav2time.csv")
xyplot(fert_sq~time,data=dat,type=c("p","r"), main = "OFAV 2019 Night 2: Time vs. Fert - SQ corrected avg (n=3)", xlab="Time Ova Added to Sperm", ylab="Fert Avg (n=3): Square Corrected")
m <- lm(fert_sq~time,data=dat)
anova(m)
```
```{r}
summary(m)
```

iii. 2022 Night 1

```{r}
library(lattice)
rm(list=ls())
dat <- read.csv("C:/Users/parsonsee/Desktop/ofav4time.csv")
xyplot(fert_sq~time,data=dat,type=c("p","r"), main = "OFAV 2022: Time vs. Fert - SQ corrected avg (n=3)", xlab="Time Ova Added to Sperm", ylab="Fert Avg (n=3): Square Corrected")
m <- lm(fert_sq~time,data=dat)
anova(m)
```

```{r}
summary(m)
```

iv. 2022 Night 2

```{r}
library(lattice)
rm(list=ls())
dat <- read.csv("C:/Users/parsonsee/Desktop/ofavtimeall.csv")
xyplot(fert~time,data=dat,type=c("p","r"), main = "All Nights: Time vs. Fertilization", ylab="Fertilization", xlab="Time (Hours) Between Spawning and Crosses")
m <- lm(fert~time,data=dat)
anova(m)
```
```{r}
summary(m)
```



2. Age vs. Fert - average fertilization rates and Square and Square root Corrected 

```{r}
library(lattice)
rm(list=ls())
dat <- read.csv("C:/Users/parsonsee/Desktop/2019night1sq.csv")
xyplot(fert_sq~time,data=dat,type=c("p","r"), main = "2019 Night 1: Time vs. Fert - SQ corrected avg (n=3)", xlab="Time Ova Added to Sperm", ylab="Fert Avg (n=3): Square Corrected") 
m <- lm(fert_sq~time,data=dat)
summary(m)
anova(m)
```


```{r}
library(lattice)
rm(list=ls())
dat <- read.csv("C:/Users/parsonsee/Desktop/2019night2sq.csv")
xyplot(fert_sq~time,data=dat,type=c("p","r"), main = "2019 Night 2: Time vs. Fert - SQ corrected avg (n=3)", xlab="Time Ova Added to Sperm", ylab="Fert Avg (n=3): Square Corrected")
m <- lm(fert_sq~time,data=dat)
summary(m)
anova(m)
```




```{r}
library(lattice)
rm(list=ls())
dat <- read.csv("C:/Users/parsonsee/Desktop/2022night1sqrt.csv")
xyplot(fert_sqrt~time,data=dat,type=c("p","r"), main = "2022 Night 1: Time vs. Fert - Square Root Corrected Avg (n=3)", xlab="Time Ova Added to Sperm", ylab="Fert Avg (n=3): Sqrt Corrected")
m <- lm(fert_sqrt~time,data=dat)
summary(m)
anova(m)
```

```{r}
library(lattice)
rm(list=ls())
dat <- read.csv("C:/Users/parsonsee/Desktop/2022night2sqrt.csv")
xyplot(fert_sqrt~time,data=dat,type=c("p","r"), main = "2022 Night 2: Time vs. Fert - Square Root Corrected Avg (n=3)", xlab="Time Ova Added to Sperm", ylab="Fert Avg (n=3): Sqrt Corrected")
m <- lm(fert_sqrt~time,data=dat)
summary(m)
anova(m)
```


3. Mixed effects of age, directionality, and genetic relatedness 




2. night 1 

```{r}
library(lattice)
rm(list=ls())
dat <- read.csv("C:/Users/parsonsee/Desktop/2019night1sq.csv")
dat$parent1 <- unlist(lapply(strsplit(dat$cross,"-"),"[[",1))
dat$parent2 <- unlist(lapply(strsplit(dat$cross,"-"),"[[",2))
xyplot(fert_sq~time*lynchli,data=dat,type=c("p","r"))
m <- lm(fert_sq~time*lynchli,data=dat)
summary(m)
anova(m)
```


```{r}
model <- lm(fert_sq ~ homog, data = dat)
summary(model)
```




Parent 1 (sperm) effects 

```{r}
plot(x=dat$homog,y=dat$fert_sq,type="n")
unique1 <- unique(dat$parent1)
stats <- c()
for(i in 1:4)
{
  tmp <- dat[dat$parent1==unique1[i],]
  m <- lm(fert_sq~homog,data=tmp)
  abline(m)
  stats <- rbind(stats,data.frame(parent1=unique1[i],n=dim(tmp)[1],slope=summary(m)$coefficients[2],r2=round(summary(m)$r.squared,3),p=round(anova(m)$"Pr(>F)"[1],3)))
}
stats$signif=ifelse(stats$p<0.05,"*","")
print(stats)
```


parent 2 (ova) effects

```{r}
plot(x=dat$homog,y=dat$fert_sq,type="n")
unique2 <- unique(dat$parent2)
stats2 <- c()
for(i in 1:4)
{
  tmp <- dat[dat$parent2==unique1[i],]
  m <- lm(fert_sq~homog,data=tmp)
  abline(m)
  stats2 <- rbind(stats2,data.frame(parent2=unique1[i],n=dim(tmp)[1],slope=summary(m)$coefficients[2],r2=round(summary(m)$r.squared,3),p=round(anova(m)$"Pr(>F)"[1],3)))
}
stats2$signif=ifelse(stats2$p<0.05,"*","")
print(stats2)
```


2019 night 2

```{r}
library(lattice)
rm(list=ls())
dat <- read.csv("C:/Users/parsonsee/Desktop/2019night2sq.csv")
dat$parent1 <- unlist(lapply(strsplit(dat$cross,"-"),"[[",1))
dat$parent2 <- unlist(lapply(strsplit(dat$cross,"-"),"[[",2))
xyplot(fert_sq~lynchli*time,data=dat,type=c("p","r"))
m <- lm(fert_sq~lynchli*time,data=dat)
summary(m)
anova(m)
```
Residual standard error
```{r}
sigma(m)/mean(dat$fert_sq)
```

Parent 1 effects 

```{r}
plot(x=dat$homog,y=dat$fert_sq,type="n")
unique1 <- unique(dat$parent1)
stats <- c()
for(i in 1:8)
{
  tmp <- dat[dat$parent1==unique1[i],]
  m <- lm(fert_sq~homog,data=tmp)
  abline(m)
  stats <- rbind(stats,data.frame(parent1=unique1[i],n=dim(tmp)[1],slope=summary(m)$coefficients[2],r2=round(summary(m)$r.squared,3),p=round(anova(m)$"Pr(>F)"[1],3)))
}
stats$signif=ifelse(stats$p<0.05,"*","")
print(stats)
```
 


## Parent 2 effects that are significant all have positive slopes

```{r}
plot(x=dat$homog,y=dat$fert_sq,type="n")
unique2 <- unique(dat$parent2)
stats2 <- c()
for(i in 1:8)
{
  tmp <- dat[dat$parent2==unique1[i],]
  m <- lm(fert_sq~homog,data=tmp)
  abline(m)
  stats2 <- rbind(stats2,data.frame(parent2=unique1[i],n=dim(tmp)[1],slope=summary(m)$coefficients[2],r2=round(summary(m)$r.squared,3),p=round(anova(m)$"Pr(>F)"[1],3)))
}
stats2$signif=ifelse(stats2$p<0.05,"*","")
print(stats2)
```



Parent 1 effect overall 

```{r}
  tmp <- dat[dat$parent1==unique1[i],]
  m <- lm(fert_sq~homog,data=tmp)
  anova(m)
```

parent 2 effect...

```{r}
  tmp <- dat[dat$parent2==unique1[i],]
  m <- lm(fert_sq~homog,data=tmp)
  anova(m)
```



4. Average - night 1

```{r}
dat <- read.csv("C:/Users/parsonsee/Desktop/night.1.csv")
mean.fert <- tapply(dat$fert,dat$cross,mean)
dat$parent1 <- unlist(lapply(strsplit(dat$cross,"-"),"[[",1))
dat$parent2 <- unlist(lapply(strsplit(dat$cross,"-"),"[[",2))
prov_R <- dat$prov_R[match(names(mean.fert),dat$cross)]
xyplot(mean.fert~prov_R,type=c("p","r"))
m <- lm(mean.fert~prov_R)
anova(m)
cor.test(x=mean.fert,y=prov_R,method="kendall")
```


Night 1 parental affects



## Parent 1 had a signigicant negatice slope for genet U50 

```{r}
plot(x=dat$prov_R,y=dat$fert,type="n")
unique1 <- unique(dat$parent1)
stats <- c()
for(i in 1:4)
{
  tmp <- dat[dat$parent1==unique1[i],]
  m <- lm(fert~prov_R,data=tmp)
  abline(m)
  stats <- rbind(stats,data.frame(parent1=unique1[i],n=dim(tmp)[1],slope=summary(m)$coefficients[2],r2=round(summary(m)$r.squared,3),p=round(anova(m)$"Pr(>F)"[1],3)))
}
stats$signif=ifelse(stats$p<0.05,"*","")
print(stats)
```



## Parent 2 - none are significant by themselves? 

```{r}
plot(x=dat$prov_R,y=dat$fert,type="n")
unique2 <- unique(dat$parent2)
stats2 <- c()
for(i in 1:4)
{
  tmp <- dat[dat$parent2==unique1[i],]
  m <- lm(fert~prov_R,data=tmp)
  abline(m)
  stats2 <- rbind(stats2,data.frame(parent2=unique1[i],n=dim(tmp)[1],slope=summary(m)$coefficients[2],r2=round(summary(m)$r.squared,3),p=round(anova(m)$"Pr(>F)"[1],3)))
}
stats2$signif=ifelse(stats2$p<0.05,"*","")
print(stats2)
```


Parent 1 effect overall 

```{r}
  tmp <- dat[dat$parent1==unique1[i],]
  m <- lm(fert~prov_R,data=tmp)
  anova(m)
```

parent 2 effect...

```{r}
  tmp <- dat[dat$parent2==unique1[i],]
  m <- lm(fert~prov_R,data=tmp)
  anova(m)
```

## When you ombine these parent1 or parent 2 effects, only 1 slope is significant - shows negative relationship.

```{r}
plot(x=dat$prov_R,y=dat$fert,type="n")
unique1 <- unique(dat$parent1)
stats3 <- c()
for(i in 1:4)
{
  tmp <- dat[dat$parent1==unique1[i] | dat$parent2==unique1[i],]
  m <- lm(fert~prov_R,data=tmp)
  abline(m)
  stats3 <- rbind(stats3,data.frame(parent=unique1[i],n=dim(tmp)[1],slope=summary(m)$coefficients[2],r2=round(summary(m)$r.squared,3),p=round(anova(m)$"Pr(>F)"[1],3)))
  
}
stats3$signif=ifelse(stats3$p<0.05,"*","")
print(stats3)
```





5. Average - night 2 


```{r}
dat <- read.csv("C:/Users/parsonsee/Desktop/night.2.csv")
mean.fert <- tapply(dat$fert,dat$cross,mean)
dat$parent1 <- unlist(lapply(strsplit(dat$cross,"-"),"[[",1))
dat$parent2 <- unlist(lapply(strsplit(dat$cross,"-"),"[[",2))
prov_R <- dat$prov_R[match(names(mean.fert),dat$cross)]
xyplot(mean.fert~prov_R,type=c("p","r"))
m <- lm(mean.fert~prov_R)
anova(m)
cor.test(x=mean.fert,y=prov_R,method="kendall")
```


All nights Average 

```{r}
library(lattice)
rm(list=ls())
dat <- read.csv("C:/Users/parsonsee/Desktop/all_time.csv")
xyplot(Fert~Time,data=dat,type=c("p","r"), main = "Time vs. Fert - All Nights", xlab="Time between Spawn and Cross", ylab="Overall Fertilization Average")
m <- lm(Fert~Time,data=dat)
summary(m)
anova(m)
```


